---
title: "Predicting Online Purchases"
subtitle: "An e-commerce conversion analytics proposal"
author: "Angela Qian"
date: today
format:
  revealjs:
    theme: [default, Angela-Qian-Online-Shoppers-Deck.scss]
    transition: fade
    slide-number: c/t
    progress: true
    center: false
    width: 1600
    height: 900
    margin: 0.06
    embed-resources: true
    auto-stretch: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(scales)

shop <- read_csv("Online-Shoppers-UCI.csv", show_col_types = FALSE) |>
  mutate(
    Revenue = factor(Revenue, levels = c(FALSE, TRUE),
                     labels = c("No purchase", "Purchase")),
    Weekend = factor(Weekend, levels = c(FALSE, TRUE),
                     labels = c("Weekday", "Weekend"))
  )
```

## The question is simple {.statement-slide}

> Which website visits are most likely to end in a purchase?

**Project goal:** build a classification model that helps an e-commerce team recognize high-conversion sessions and improve the customer journey.

::: {.notes}
[Sources]
- UCI Machine Learning Repository, Online Shoppers Purchasing Intention Dataset: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## This approved dataset is ready for analysis

:::: {.columns}
::: {.column width="57%"}

### Online Shoppers Purchasing Intention

- **Source:** UCI Machine Learning Repository
- **Subject area:** Business
- **Task:** Classification and clustering
- **Observations:** 12,330 website sessions
- **Predictors:** 17 browsing and visitor variables
- **Outcome:** Revenue, true or false
- **Missing values:** None
- **License:** CC BY 4.0

:::
::: {.column width="39%"}

```{r}
#| label: fig-outcomes
#| fig-width: 6
#| fig-height: 5
#| fig-cap: "Observed session outcomes"
shop |>
  count(Revenue) |>
  mutate(share = n / sum(n)) |>
  ggplot(aes(Revenue, share, fill = Revenue)) +
  geom_col(width = .62, show.legend = FALSE) +
  geom_text(aes(label = percent(share, accuracy = .1)),
            vjust = -0.45, fontface = "bold", size = 7) +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 18) +
  theme(panel.grid = element_blank())
```

:::
::::

::: {.notes}
[Sources]
- UCI dataset documentation: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
- Dataset DOI: https://doi.org/10.24432/C5F88Q
:::

## The variables describe a familiar shopping journey

:::: {.columns}
::: {.column width="48%"}

### What visitors did

- Administrative pages and time
- Informational pages and time
- Product pages and time
- Bounce rate and exit rate
- Page value

:::
::: {.column width="48%"}

### Who and when

- New or returning visitor
- Traffic type and region
- Browser and operating system
- Month and weekend
- Closeness to a special day

:::
::::

The target, **Revenue**, records whether the session ended in a transaction.

::: {.notes}
[Sources]
- UCI variable definitions: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The business problem is lost conversion opportunity

Most sessions do not end in a purchase. The e-commerce team needs to understand where purchasing behavior differs from non-purchasing behavior.

:::: {.columns}
::: {.column width="31%"}

### Diagnose

Find behaviors connected with successful visits.

:::
::: {.column width="31%"}

### Predict

Estimate the chance that a session will produce revenue.

:::
::: {.column width="31%"}

### Improve

Use the findings to guide site and campaign decisions.

:::
::::

## The e-commerce manager is the main audience

The **E-commerce Marketing Manager** needs clear evidence about which sessions are more likely to convert and where customers may be dropping out.

The findings can also help:

- Website managers improve product-page journeys
- Paid media teams compare traffic quality
- CRM teams plan remarketing for visitors who do not purchase
- Executives monitor conversion opportunities

## Three questions keep the project focused

1. **Prediction:** How accurately can a purchase be predicted from session behavior?
2. **Drivers:** Which browsing and visitor variables are most useful for distinguishing purchases?
3. **Action:** Which customer journeys or traffic segments should receive attention first?

The outcome is binary, making logistic regression a clear baseline that is easy to explain.

## Purchasing sessions spend more time with products

```{r}
#| label: fig-product-time
#| fig-width: 11
#| fig-height: 4.3
#| fig-cap: "Median product-related duration by session outcome"
shop |>
  group_by(Revenue) |>
  summarise(`Median seconds` = median(ProductRelated_Duration), .groups = "drop") |>
  ggplot(aes(Revenue, `Median seconds`, fill = Revenue)) +
  geom_col(width = .58, show.legend = FALSE) +
  geom_text(aes(label = comma(`Median seconds`, accuracy = 1)),
            vjust = -0.45, size = 7, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = comma, limits = c(0, 1300)) +
  labs(x = NULL, y = "Median product-related duration (seconds)") +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.x = element_blank())
```

Purchasing sessions spent about twice as long on product pages. This is a pattern, not proof of cause.

::: {.notes}
[Sources]
- Calculations use online_shoppers_intention.csv from UCI: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## Conversion changes across the year

```{r}
#| label: fig-month
#| fig-width: 11
#| fig-height: 4.4
#| fig-cap: "Observed purchase rate by month"
month_order <- c("Feb", "Mar", "May", "June", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

shop |>
  mutate(Month = factor(Month, levels = month_order)) |>
  group_by(Month) |>
  summarise(rate = mean(Revenue == "Purchase"), .groups = "drop") |>
  ggplot(aes(Month, rate, group = 1)) +
  geom_line(color = "#326B77", linewidth = 1.6) +
  geom_point(aes(color = rate), size = 5) +
  scale_color_gradient(low = "#8EC6C5", high = "#F28E5B", guide = "none") +
  scale_y_continuous(labels = percent, limits = c(0, .30)) +
  labs(x = NULL, y = "Observed purchase rate") +
  theme_minimal(base_size = 18) +
  theme(panel.grid.minor = element_blank())
```

November had the highest purchase rate (25.4%), so timing should be tested in the model.

::: {.notes}
[Sources]
- Calculations use online_shoppers_intention.csv from UCI: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The analysis can stay practical and manageable

:::: {.columns}
::: {.column width="47%"}

### Modeling steps

1. Check distributions and class balance
2. Create a stratified train/test split
3. Prepare categorical predictors
4. Fit logistic regression
5. Compare one tree-based model
6. Evaluate on the test set

:::
::: {.column width="49%"}

### Evaluation measures

- **Recall:** purchases correctly identified
- **Precision:** how often a purchase prediction is right
- **ROC AUC:** overall ranking quality
- **Confusion matrix:** correct and incorrect decisions

Because only 15.5% of sessions purchased, accuracy alone will not be enough.

:::
::::

## The project can lead to clear marketing actions

If the model performs well, the company can:

- Identify the browsing patterns most connected with purchases
- Improve pages where visitors commonly exit
- Compare the quality of traffic sources
- Time campaigns around stronger conversion periods
- Build better remarketing audiences

**Recommendation:** approve the Online Shoppers dataset and begin with exploratory analysis and logistic regression.

## Approved source

**Dataset:** Online Shoppers Purchasing Intention Dataset  
<https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset>

**DOI:** <https://doi.org/10.24432/C5F88Q>

**License:** Creative Commons Attribution 4.0 International

**Downloaded file:** `Online-Shoppers-UCI.csv`

This project offers a direct business question, understandable variables, and a simple outcome that is easy to present.

::: {.notes}
[Sources]
- UCI Machine Learning Repository: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
- CC BY 4.0 license: https://creativecommons.org/licenses/by/4.0/legalcode
:::
