---
title: "Predicting Online Purchases"
subtitle: "Turning website behavior into conversion decisions"
author:
  - Angela Qian
  - Bhamini Pillai
date: today
format:
  revealjs:
    theme: [default, Angela-Qian-Online-Shoppers-Final.scss]
    transition: fade
    slide-number: c/t
    progress: true
    center: false
    width: 1600
    height: 900
    margin: 0.055
    embed-resources: true
    auto-stretch: false
    code-overflow: wrap
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
#| label: setup
library(tidymodels)
library(tidyverse)
library(scales)

set.seed(6540)

shop <- read_csv("Online-Shoppers-UCI.csv", show_col_types = FALSE) |>
  mutate(
    Revenue = factor(Revenue, levels = c(FALSE, TRUE),
                     labels = c("No purchase", "Purchase")),
    Weekend = factor(Weekend, levels = c(FALSE, TRUE),
                     labels = c("Weekday", "Weekend")),
    across(c(OperatingSystems, Browser, Region, TrafficType), factor)
  )

shop_split <- initial_split(shop, prop = .80, strata = Revenue)
shop_train <- training(shop_split)
shop_test <- testing(shop_split)

shop_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

log_spec <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification")

tree_spec <- decision_tree(cost_complexity = .002, tree_depth = 8, min_n = 20) |>
  set_engine("rpart") |>
  set_mode("classification")

rf_spec <- rand_forest(mtry = 7, trees = 500, min_n = 12) |>
  set_engine("ranger", importance = "permutation", probability = TRUE) |>
  set_mode("classification")

fit_model <- function(spec) {
  workflow() |>
    add_recipe(shop_rec) |>
    add_model(spec) |>
    fit(shop_train)
}

log_fit <- fit_model(log_spec)
tree_fit <- fit_model(tree_spec)
rf_fit <- fit_model(rf_spec)

score_model <- function(fit, label) {
  scored <- bind_cols(
    shop_test |> select(Revenue),
    predict(fit, shop_test, type = "prob"),
    predict(fit, shop_test, type = "class")
  )
  tibble(
    Model = label,
    `ROC AUC` = roc_auc_vec(scored$Revenue, scored$.pred_Purchase,
                            event_level = "second"),
    Accuracy = accuracy_vec(scored$Revenue, scored$.pred_class),
    Recall = sens_vec(scored$Revenue, scored$.pred_class,
                      event_level = "second"),
    Precision = precision_vec(scored$Revenue, scored$.pred_class,
                              event_level = "second")
  )
}

model_metrics <- bind_rows(
  score_model(log_fit, "Logistic regression"),
  score_model(tree_fit, "Decision tree"),
  score_model(rf_fit, "Random forest")
)

rf_pred <- bind_cols(
  shop_test |> select(Revenue),
  predict(rf_fit, shop_test, type = "prob")
) |>
  mutate(
    pred_50 = factor(if_else(.pred_Purchase >= .50, "Purchase", "No purchase"),
                     levels = levels(Revenue)),
    pred_30 = factor(if_else(.pred_Purchase >= .30, "Purchase", "No purchase"),
                     levels = levels(Revenue))
  )

threshold_metrics <- tibble(
  Threshold = c("50%", "30%"),
  Recall = c(
    sens_vec(rf_pred$Revenue, rf_pred$pred_50, event_level = "second"),
    sens_vec(rf_pred$Revenue, rf_pred$pred_30, event_level = "second")
  ),
  Precision = c(
    precision_vec(rf_pred$Revenue, rf_pred$pred_50, event_level = "second"),
    precision_vec(rf_pred$Revenue, rf_pred$pred_30, event_level = "second")
  )
)

rf_importance <- enframe(
  extract_fit_engine(rf_fit)$variable.importance,
  name = "Feature", value = "Importance"
) |>
  arrange(desc(Importance)) |>
  slice_head(n = 10) |>
  mutate(Feature = str_replace_all(Feature, "_", " "))

no_page_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_rm(PageValues) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

no_page_fit <- workflow() |>
  add_recipe(no_page_rec) |>
  add_model(log_spec) |>
  fit(shop_train)

no_page_metrics <- score_model(no_page_fit, "Logistic model without page value")

duration_grid <- quantile(shop_train$ProductRelated_Duration,
                          probs = seq(.05, .95, length.out = 15))
duration_pdp <- map_dfr(duration_grid, function(x) {
  new_data <- shop_test
  new_data$ProductRelated_Duration <- x
  tibble(
    Duration = x,
    Probability = mean(predict(no_page_fit, new_data, type = "prob")$.pred_Purchase)
  )
})

month_order <- c("Feb", "Mar", "May", "June", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month_pdp <- map_dfr(month_order, function(x) {
  new_data <- shop_test
  new_data$Month <- x
  tibble(Month = x,
         Probability = mean(predict(rf_fit, new_data, type = "prob")$.pred_Purchase))
}) |>
  mutate(Month = factor(Month, levels = month_order))

visitor_pdp <- map_dfr(unique(shop_train$VisitorType), function(x) {
  new_data <- shop_test
  new_data$VisitorType <- x
  tibble(
    VisitorType = str_replace_all(x, "_", " "),
    Probability = mean(predict(rf_fit, new_data, type = "prob")$.pred_Purchase)
  )
}) |>
  arrange(desc(Probability))

weekend_observed <- shop |>
  group_by(Weekend) |>
  summarise(
    PurchaseRate = mean(Revenue == "Purchase"),
    Sessions = n(),
    Purchases = sum(Revenue == "Purchase"),
    .groups = "drop"
  )

weekend_pdp <- map_dfr(levels(shop_train$Weekend), function(x) {
  new_data <- shop_test
  new_data$Weekend <- factor(x, levels = levels(shop_train$Weekend))
  tibble(
    Weekend = x,
    Probability = mean(
      predict(rf_fit, new_data, type = "prob")$.pred_Purchase
    )
  )
})

test_auc <- model_metrics |> filter(Model == "Random forest") |> pull(`ROC AUC`)
```

## Where this presentation is going {.agenda-slide}

1. Business problem and data
2. AO1: Compare models
3. AO2: Explain behavioral drivers
4. AO3: Visitor and seasonal patterns
5. AO4: Weekend vs. weekday behavior
6. Recommendations and next steps

## Every session is a conversion opportunity

An online store receives thousands of visits, but most visitors leave without purchasing.

> The business challenge is to recognize promising sessions and understand where the customer journey breaks down.

Our analysis connects browsing behavior to a clear outcome: **did the session produce revenue?**

::: {.notes}
[Sources]
- UCI Machine Learning Repository: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The decision problem is practical

The e-commerce team wants to answer four questions:

- Which sessions are most likely to end in a purchase?
- Which browsing behaviors are most strongly associated with purchase outcomes?
- How do visitor type and seasonal timing relate to purchase likelihood?
- Do weekend and weekday sessions differ in purchase likelihood?

**Business goal:** use these insights to prioritize conversion opportunities, improve the customer journey, and guide campaign planning.

## Four objectives divide the work clearly {.team-slide}

| Owner | Analytics objective | Business outcome |
|---|---|---|
| Angela Qian | Compare classification models | Choose a reliable purchase-risk model |
| Angela Qian | Explain behavioral drivers | Identify journey signals that matter |
| Bhamini Pillai | Evaluate visitor and seasonal patterns | Improve audience and campaign timing |
| Bhamini Pillai | Evaluate weekend vs. weekday behavior | Assess day-specific opportunities |

## Conversion analysis matters beyond prediction

Better conversion knowledge can help a business:

- Spend marketing resources on higher-quality traffic
- Improve pages associated with customer exit
- Build more focused remarketing audiences
- Plan campaigns around stronger purchasing periods
- Monitor performance with repeatable model metrics

## Population and sample

:::: {.columns}
::: {.column width="52%"}

**Population**  
Online shopping sessions represented by the UCI study population.

**Sample**

- 12,330 observed sessions over one year
- Each session belongs to a different user
- 17 predictors plus the `Revenue` outcome
- No missing values reported

Purchasers spent a median **1,110 seconds** on product pages versus **510 seconds** for non-purchasers.

:::
::: {.column width="44%"}

```{r}
#| label: fig-class-balance-final
#| fig-width: 5.5
#| fig-height: 4.2
shop |>
  count(Revenue) |>
  mutate(Share = n / sum(n)) |>
  ggplot(aes(Revenue, Share, fill = Revenue)) +
  geom_col(width = .62, show.legend = FALSE) +
  geom_text(aes(label = percent(Share, accuracy = .1)),
            vjust = -.4, fontface = "bold", size = 6.5) +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 17) +
  theme(panel.grid = element_blank())
```

:::
::::

::: {.notes}
[Sources]
- UCI dataset documentation: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The predictors follow the customer journey

:::: {.columns}
::: {.column width="31%"}

### Engagement

Page counts and time spent on administrative, informational, and product pages

:::
::: {.column width="31%"}

### Friction and value

Bounce rate, exit rate, page value, and proximity to a special day

:::
::: {.column width="31%"}

### Context

Month, visitor type, weekend, traffic type, browser, region, and operating system

:::
::::

::: {.notes}
[Sources]
- UCI variable definitions: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## Wrangling made the data model-ready

```{r}
#| echo: true
#| eval: false
shop <- read_csv("Online-Shoppers-UCI.csv") |>
  mutate(
    Revenue = factor(Revenue,
      levels = c(FALSE, TRUE),
      labels = c("No purchase", "Purchase")),
    Weekend = factor(Weekend),
    across(c(OperatingSystems, Browser,
             Region, TrafficType), factor)
  )
```

We checked variable types, missingness, class balance, and plausible ranges before modeling.

## Objective 1: Compare classification models {.section-slide}

**Owner: Angela Qian**

Compare classification models to identify the most reliable approach for ranking sessions by purchase likelihood.

## We tested three classification approaches

| Method | Why include it? |
|---|---|
| Logistic regression | Clear baseline and interpretable coefficients |
| Decision tree | Simple nonlinear rules and interactions |
| Random forest | Ensemble that can capture complex patterns |

Regression-focused methods were not selected because the business outcome is categorical.

## The split protects the final evaluation

```{r}
#| echo: true
#| eval: false
shop_split <- initial_split(
  shop, prop = .80, strata = Revenue
)

shop_train <- training(shop_split)
shop_test  <- testing(shop_split)
```

- **Training set:** 80% of sessions
- **Test set:** 20% of sessions
- **Stratification:** maintains approximately the 15.5% purchase-class proportion in both sets

## One recipe keeps preprocessing consistent

```{r}
#| echo: true
#| eval: false
shop_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())
```

The same recipe was used for all three candidate models.

## The random forest ranked sessions best

```{r}
#| label: fig-model-comparison
#| fig-width: 10.5
#| fig-height: 4.5
model_metrics |>
  select(Model, `ROC AUC`, Recall, Precision) |>
  pivot_longer(-Model, names_to = "Metric", values_to = "Value") |>
  ggplot(aes(Value, Model, color = Metric)) +
  geom_point(size = 6) +
  scale_color_manual(values = c("#326B77", "#E9A23B", "#F28E5B")) +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top", panel.grid.major.y = element_blank())
```

The random forest achieved **`r number(test_auc, accuracy = .001)` ROC AUC** on unseen test data.

## A lower threshold finds more purchasers

```{r}
#| label: fig-thresholds
#| fig-width: 10
#| fig-height: 4.2
threshold_metrics |>
  pivot_longer(-Threshold, names_to = "Metric", values_to = "Value") |>
  ggplot(aes(Threshold, Value, fill = Metric)) +
  geom_col(position = position_dodge(.75), width = .68) +
  geom_text(aes(label = percent(Value, accuracy = .1)),
            position = position_dodge(.75), vjust = -.35,
            size = 5.7, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, .9)) +
  labs(x = "Purchase probability threshold", y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top", panel.grid.major.x = element_blank())
```

At 30%, recall rises to 75.1% while precision remains 62.5%.

## Objective 1 finding {.finding-slide}

> Use the random forest for ranking sessions, then select the threshold based on outreach capacity and the cost of missed purchasers.

- Best test ROC AUC: 0.919
- A 30% threshold finds three out of four purchasing sessions
- A 50% threshold produces fewer but more precise alerts

## Objective 2: What behavior signals conversion? {.section-slide}

**Owner: Angela Qian**

Explain which customer-journey variables matter and whether engagement patterns remain useful without the strongest page-value signal.

## Purchasers spend longer on product pages

```{r}
#| label: fig-duration-final
#| fig-width: 9.5
#| fig-height: 4.3
shop |>
  group_by(Revenue) |>
  summarise(Duration = median(ProductRelated_Duration), .groups = "drop") |>
  ggplot(aes(Revenue, Duration, fill = Revenue)) +
  geom_col(width = .58, show.legend = FALSE) +
  geom_text(aes(label = comma(Duration, accuracy = 1)),
            vjust = -.4, size = 6.2, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(limits = c(0, 1300), labels = comma) +
  labs(x = NULL, y = "Median product-page duration (seconds)") +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.x = element_blank())
```

Purchasing sessions had a median product-page duration of 1,110 seconds versus 510 seconds.

## PageValue dominates model importance

```{r}
#| label: fig-importance
#| fig-width: 10
#| fig-height: 4.6
rf_importance |>
  slice_head(n = 8) |>
  ggplot(aes(Importance, reorder(Feature, Importance))) +
  geom_col(fill = "#326B77", width = .68) +
  labs(x = "Permutation importance", y = NULL) +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.y = element_blank())
```

`PageValues` is the strongest predictor in the model.

Other useful signals include exit behavior, product engagement, and administrative activity. Because `PageValues` is closely connected to transaction activity, interpret it cautiously as an early behavioral signal.

## Can browsing behavior predict purchase without PageValue?

**Why remove PageValue?** It is closely connected with transaction activity, so we fitted a second logistic regression without it.

```{r}
#| echo: true
#| eval: false
behavior_recipe <- recipe(Revenue ~ ., data = shop_train) |>
  step_rm(PageValues) |>
  step_dummy(all_nominal_predictors()) |>
  step_normalize(all_numeric_predictors())
```

**Purpose:** determine whether earlier browsing behavior still contains useful predictive information.

This is a sensitivity analysis—not a replacement for the main model.

## Behavior alone provides moderate ranking power

:::: {.columns}
::: {.column width="48%"}

### Without page value

- ROC AUC: **0.715**
- Accuracy: **83.9%**
- Recall at 50%: **3.7%**

:::
::: {.column width="48%"}

### Interpretation

Browsing behavior provides useful ranking information, but the default 50% threshold misses most purchasers.

Because only 15.5% of sessions result in a purchase, accuracy alone can be misleading.

:::
::::

## Longer product engagement raises predicted probability

```{r}
#| label: fig-duration-pdp
#| fig-width: 10
#| fig-height: 4.4
duration_pdp |>
  ggplot(aes(Duration, Probability)) +
  geom_line(color = "#326B77", linewidth = 1.7) +
  geom_point(color = "#F28E5B", size = 3.5) +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = percent) +
  labs(x = "Product-related duration (seconds)",
       y = "Average predicted purchase probability") +
  theme_minimal(base_size = 18)
```

The partial-dependence pattern is an association, not a causal effect.

## Objective 2 finding {.finding-slide}

> Page value is the strongest model signal, while product engagement and exit behavior provide earlier clues about conversion.

- Use page value carefully because it is close to the transaction outcome
- Monitor exit rate as a possible friction signal
- Test assistance or remarketing for sessions showing sustained product interest

## Objective 3: Visitor and seasonal patterns {.section-slide}

**Owner: Bhamini Pillai**

Estimate how purchase probability changes across months and visitor types while other recorded characteristics remain constant.

## Observed session-level purchase rate peaks in November

```{r}
#| label: fig-observed-month-final
#| fig-width: 10.5
#| fig-height: 4.4
shop |>
  mutate(Month = factor(Month, levels = month_order)) |>
  group_by(Month) |>
  summarise(Rate = mean(Revenue == "Purchase"), .groups = "drop") |>
  ggplot(aes(Month, Rate, group = 1)) +
  geom_line(color = "#326B77", linewidth = 1.6) +
  geom_point(aes(color = Rate), size = 4.8) +
  scale_color_gradient(low = "#8EC6C5", high = "#F28E5B", guide = "none") +
  scale_y_continuous(labels = percent, limits = c(0, .30)) +
  labs(x = NULL, y = "Observed purchase rate") +
  theme_minimal(base_size = 18)
```

**November: 25.4%**

November shows a stronger purchasing period in this dataset and may warrant additional campaign attention.

## The segment analysis uses the ensemble model

For every month or visitor type, we:

1. Set that feature to one value for all test sessions
2. Predicted each session with the fitted random forest
3. Averaged the predicted purchase probabilities
4. Repeated the process for every category

This partial-dependence approach summarizes model behavior while holding the test population constant.

## Month remains important after adjustment

```{r}
#| label: fig-month-pdp-final
#| fig-width: 10.5
#| fig-height: 4.4
month_pdp |>
  ggplot(aes(Month, Probability, group = 1)) +
  geom_line(color = "#326B77", linewidth = 1.6) +
  geom_point(color = "#F28E5B", size = 4.8) +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Average predicted purchase probability") +
  theme_minimal(base_size = 18)
```

The model assigns the strongest adjusted probabilities to late-year sessions.

## Visitor type changes predicted conversion

```{r}
#| label: fig-visitor-pdp-final
#| fig-width: 9.5
#| fig-height: 4.3
visitor_pdp |>
  ggplot(aes(Probability, reorder(VisitorType, Probability))) +
  geom_col(fill = "#326B77", width = .62) +
  geom_text(aes(label = percent(Probability, accuracy = .1)),
            hjust = -0.2, fontface = "bold", size = 5.8) +
  scale_x_continuous(labels = percent, limits = c(0, .23)) +
  labs(x = "Average predicted purchase probability", y = NULL) +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.y = element_blank())
```

Visitor type should inform audience strategy, but the differences are smaller than the page-value signal.

## Objective 3 finding {.finding-slide}

> Timing and visitor type can improve campaign planning, but they should complement—not replace—session behavior.

- Prepare conversion campaigns before the late-year peak
- Test different messages for new and returning visitors
- Compare performance across visitor types and months
- Validate segment strategies through controlled experiments

## Objective 4: Weekend vs. weekday behavior {.section-slide}

**Owner: Bhamini Pillai**

Do weekend sessions have different purchase likelihood than weekday sessions?

Understanding day-of-week differences can help an e-commerce team decide whether campaign timing, promotions, remarketing, or website messaging should be tested differently.

## Weekend sessions show a higher purchase rate

```{r}
#| label: fig-weekend-observed
#| fig-width: 9.5
#| fig-height: 4.4
weekend_observed |>
  ggplot(aes(Weekend, PurchaseRate, fill = Weekend)) +
  geom_col(width = .58, show.legend = FALSE) +
  geom_text(aes(label = percent(PurchaseRate, accuracy = .1)),
            vjust = -.4, size = 6.3, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, .22)) +
  labs(x = NULL, y = "Observed purchase rate") +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.x = element_blank())
```

Weekend sessions had a **2.5 percentage-point higher** observed purchase rate than weekday sessions. This is an association—not evidence that weekends cause purchasing.

## Weekend effect as a model-based comparison

:::: {.columns}
::: {.column width="48%"}

### Analysis approach

1. Set all test sessions to weekday
2. Predict purchase probability
3. Set all test sessions to weekend
4. Predict again and compare averages

:::
::: {.column width="48%"}

```{r}
#| label: fig-weekend-pdp
#| fig-width: 5.4
#| fig-height: 4.2
weekend_pdp |>
  ggplot(aes(Weekend, Probability, fill = Weekend)) +
  geom_col(width = .58, show.legend = FALSE) +
  geom_text(aes(label = percent(Probability, accuracy = .1)),
            vjust = -.4, size = 6, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, .22)) +
  labs(x = NULL, y = "Average predicted probability") +
  theme_minimal(base_size = 16) +
  theme(panel.grid.major.x = element_blank())
```

:::
::::

This isolates the model's estimated weekend relationship while keeping the test population constant. The observed and model-based views answer different questions; neither establishes causality.

## Objective 4 finding {.finding-slide}

> Weekend timing is a promising signal to test—not a reason to automatically shift marketing spending.

- Test whether weekend campaigns generate stronger conversion
- Compare promotional messaging across weekdays and weekends
- Consider weekend-specific remarketing experiments
- Validate any timing strategy through controlled experiments

## Four findings answer the business problem {.summary-slide}

1. **Prediction:** the random forest ranked purchase likelihood well, with 0.919 test ROC AUC.
2. **Behavior:** page value led the model; product engagement and exit behavior added useful signals.
3. **Timing and visitor type:** late-year timing and visitor type were associated with different predicted purchase probabilities.
4. **Weekend behavior:** weekend sessions showed a higher observed purchase rate than weekday sessions.

**Overall:** purchase likelihood is best understood by combining session behavior with visitor and timing context.

## Recommended business actions {.recommend-slide}

:::: {.action-grid}
::: {.action-item}

### Prioritize

Score sessions and select a threshold based on capacity and business costs.

:::
::: {.action-item}

### Improve

Investigate high-exit journeys and test help for sustained product interest.

:::
::: {.action-item}

### Time

Test whether weekend and late-year campaigns produce stronger conversion.

:::
::: {.action-item}

### Test

Run controlled experiments on audience, timing, messaging, and remarketing.

:::
::::

**Bottom line:** use the model to identify opportunities, then test whether the actions improve conversion.

## Results have important limitations

- The data represents sessions, not a longitudinal customer history
- The source does not identify the retailer or product categories
- Page value may contain information close to the final transaction
- Class imbalance makes accuracy potentially misleading
- Model relationships do not prove that a business action will cause conversion
- Weekend/weekday differences are observational and require experimental validation before changing campaign allocation
- Current performance should be validated on newer company data

## Next steps move from model to decision

1. Tune hyperparameters with cross-validation
2. Compare probability calibration and business costs
3. Validate the threshold using campaign capacity
4. Test explanations with SHAP values
5. Test weekend vs. weekday campaign strategies
6. Use the deployed scoring app in a controlled pilot
7. Measure incremental conversion through an experiment

## Appendix {.appendix-slide}

::: {.appendix-card}
### Project resources

**GitHub repository**  
[View the Group 6 project repository](https://github.com/angelakew/ibm6540-final-report)

**Presentation slides**  
[Open the published RevealJS presentation](https://angelakew.github.io/ibm6540-final-report/)

**Project dataset**  
[View the UCI Online Shoppers dataset abstract](https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset)

**Dataset citation**  
[UCI dataset DOI: 10.24432/C5F88Q](https://doi.org/10.24432/C5F88Q)

**Model availability**  
The reproducible model analysis is included in the project QMD.  
[Open the scoring app](https://angelakew.shinyapps.io/online-purchase-explorer/)
:::

::: {.notes}
[Sources]
- UCI Machine Learning Repository: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
- Dataset DOI: https://doi.org/10.24432/C5F88Q
:::

## Reproducibility notes

- Random seed: 6540
- Stratified 80/20 train/test split
- Shared preprocessing recipe across candidate models
- Test set used only for final comparison
- Weekend analysis uses the same test-set prediction framework as the other model-based analyses
- R packages: tidymodels, tidyverse, ranger, and Quarto
- Complete analysis code is contained in this QMD file

## The business value is focused learning {.statement-slide .closing-slide}

> Predict purchase likelihood, explain the customer journey, and test the actions that follow.

The model is a starting point for better decisions—not a substitute for experimentation.

### Project links

[GitHub repository](https://github.com/angelakew/ibm6540-final-report) · [Published presentation](https://angelakew.github.io/ibm6540-final-report/)
