---
title: "Predicting Online Purchases"
subtitle: "Turning website behavior into conversion decisions"
author:
  - Angela Qian
  - Bhamini Pillai
  - Willy Su
date: today
format:
  revealjs:
    theme: [default, Angela-Qian-Online-Shoppers-Final.scss]
    transition: fade
    slide-number: c/t
    progress: true
    center: false
    width: 1600
    height: 900
    margin: 0.055
    embed-resources: true
    auto-stretch: false
    code-overflow: wrap
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
#| label: setup
library(tidymodels)
library(tidyverse)
library(scales)

set.seed(6540)

shop <- read_csv("Online-Shoppers-UCI.csv", show_col_types = FALSE) |>
  mutate(
    Revenue = factor(Revenue, levels = c(FALSE, TRUE),
                     labels = c("No purchase", "Purchase")),
    Weekend = factor(Weekend, levels = c(FALSE, TRUE),
                     labels = c("Weekday", "Weekend")),
    across(c(OperatingSystems, Browser, Region, TrafficType), factor)
  )

shop_split <- initial_split(shop, prop = .80, strata = Revenue)
shop_train <- training(shop_split)
shop_test <- testing(shop_split)

shop_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

log_spec <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification")

tree_spec <- decision_tree(cost_complexity = .002, tree_depth = 8, min_n = 20) |>
  set_engine("rpart") |>
  set_mode("classification")

rf_spec <- rand_forest(mtry = 7, trees = 500, min_n = 12) |>
  set_engine("ranger", importance = "permutation", probability = TRUE) |>
  set_mode("classification")

fit_model <- function(spec) {
  workflow() |>
    add_recipe(shop_rec) |>
    add_model(spec) |>
    fit(shop_train)
}

log_fit <- fit_model(log_spec)
tree_fit <- fit_model(tree_spec)
rf_fit <- fit_model(rf_spec)

score_model <- function(fit, label) {
  scored <- bind_cols(
    shop_test |> select(Revenue),
    predict(fit, shop_test, type = "prob"),
    predict(fit, shop_test, type = "class")
  )
  tibble(
    Model = label,
    `ROC AUC` = roc_auc_vec(scored$Revenue, scored$.pred_Purchase,
                            event_level = "second"),
    Accuracy = accuracy_vec(scored$Revenue, scored$.pred_class),
    Recall = sens_vec(scored$Revenue, scored$.pred_class,
                      event_level = "second"),
    Precision = precision_vec(scored$Revenue, scored$.pred_class,
                              event_level = "second")
  )
}

model_metrics <- bind_rows(
  score_model(log_fit, "Logistic regression"),
  score_model(tree_fit, "Decision tree"),
  score_model(rf_fit, "Random forest")
)

rf_pred <- bind_cols(
  shop_test |> select(Revenue),
  predict(rf_fit, shop_test, type = "prob")
) |>
  mutate(
    pred_50 = factor(if_else(.pred_Purchase >= .50, "Purchase", "No purchase"),
                     levels = levels(Revenue)),
    pred_30 = factor(if_else(.pred_Purchase >= .30, "Purchase", "No purchase"),
                     levels = levels(Revenue))
  )

threshold_metrics <- tibble(
  Threshold = c("50%", "30%"),
  Recall = c(
    sens_vec(rf_pred$Revenue, rf_pred$pred_50, event_level = "second"),
    sens_vec(rf_pred$Revenue, rf_pred$pred_30, event_level = "second")
  ),
  Precision = c(
    precision_vec(rf_pred$Revenue, rf_pred$pred_50, event_level = "second"),
    precision_vec(rf_pred$Revenue, rf_pred$pred_30, event_level = "second")
  )
)

rf_importance <- enframe(
  extract_fit_engine(rf_fit)$variable.importance,
  name = "Feature", value = "Importance"
) |>
  arrange(desc(Importance)) |>
  slice_head(n = 10) |>
  mutate(Feature = str_replace_all(Feature, "_", " "))

no_page_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_rm(PageValues) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

no_page_fit <- workflow() |>
  add_recipe(no_page_rec) |>
  add_model(log_spec) |>
  fit(shop_train)

no_page_metrics <- score_model(no_page_fit, "Logistic model without page value")

duration_grid <- quantile(shop_train$ProductRelated_Duration,
                          probs = seq(.05, .95, length.out = 15))
duration_pdp <- map_dfr(duration_grid, function(x) {
  new_data <- shop_test
  new_data$ProductRelated_Duration <- x
  tibble(
    Duration = x,
    Probability = mean(predict(no_page_fit, new_data, type = "prob")$.pred_Purchase)
  )
})

month_order <- c("Feb", "Mar", "May", "June", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month_pdp <- map_dfr(month_order, function(x) {
  new_data <- shop_test
  new_data$Month <- x
  tibble(Month = x,
         Probability = mean(predict(rf_fit, new_data, type = "prob")$.pred_Purchase))
}) |>
  mutate(Month = factor(Month, levels = month_order))

visitor_pdp <- map_dfr(unique(shop_train$VisitorType), function(x) {
  new_data <- shop_test
  new_data$VisitorType <- x
  tibble(
    VisitorType = str_replace_all(x, "_", " "),
    Probability = mean(predict(rf_fit, new_data, type = "prob")$.pred_Purchase)
  )
}) |>
  arrange(desc(Probability))

test_auc <- model_metrics |> filter(Model == "Random forest") |> pull(`ROC AUC`)
```

## Where this presentation is going {.agenda-slide}

1. Business problem and data
2. Objective 1: predict purchases
3. Objective 2: explain conversion behavior
4. Objective 3: find audience and timing opportunities
5. Recommendations, limitations, and next steps

## Every session is a conversion opportunity

An online store receives thousands of visits, but most visitors leave without purchasing.

> The business challenge is to recognize promising sessions and understand where the customer journey breaks down.

Our analysis connects browsing behavior to a clear outcome: **did the session produce revenue?**

::: {.notes}
[Sources]
- UCI Machine Learning Repository: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The decision problem is practical

The e-commerce team wants to answer three questions:

- Which sessions are most likely to end in a purchase?
- Which behaviors best separate purchases from non-purchases?
- Which visitor and seasonal segments offer the strongest opportunities?

These answers can guide website improvements, media spending, and remarketing.

## Three objectives divide the work clearly {.team-slide}

| Owner | Analytics objective | Business outcome |
|---|---|---|
| Angela Qian | Compare classification models | Choose a reliable purchase-risk model |
| Bhamini Pillai | Explain behavioral drivers | Identify journey signals that matter |
| Willy Su | Evaluate visitor and seasonal patterns | Improve audience and campaign timing |

## Conversion analysis matters beyond prediction

Better conversion knowledge can help a business:

- Spend marketing resources on higher-quality traffic
- Improve pages associated with customer exit
- Build more focused remarketing audiences
- Plan campaigns around stronger purchasing periods
- Monitor performance with repeatable model metrics

## The sample represents 12,330 sessions

:::: {.columns}
::: {.column width="52%"}

- One year of e-commerce sessions
- Each record represents a different user session
- 17 browsing, visitor, device, and timing predictors
- Binary outcome: purchase or no purchase
- No reported missing values
- Business dataset licensed under CC BY 4.0

:::
::: {.column width="44%"}

```{r}
#| label: fig-class-balance-final
#| fig-width: 5.5
#| fig-height: 4.2
shop |>
  count(Revenue) |>
  mutate(Share = n / sum(n)) |>
  ggplot(aes(Revenue, Share, fill = Revenue)) +
  geom_col(width = .62, show.legend = FALSE) +
  geom_text(aes(label = percent(Share, accuracy = .1)),
            vjust = -.4, fontface = "bold", size = 6.5) +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 17) +
  theme(panel.grid = element_blank())
```

:::
::::

::: {.notes}
[Sources]
- UCI dataset documentation: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## The predictors follow the customer journey

:::: {.columns}
::: {.column width="31%"}

### Engagement

Page counts and time spent on administrative, informational, and product pages

:::
::: {.column width="31%"}

### Friction and value

Bounce rate, exit rate, page value, and proximity to a special day

:::
::: {.column width="31%"}

### Context

Month, visitor type, weekend, traffic source, browser, region, and operating system

:::
::::

::: {.notes}
[Sources]
- UCI variable definitions: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
:::

## Wrangling made the data model-ready

```{r}
#| echo: true
#| eval: false
shop <- read_csv("Online-Shoppers-UCI.csv") |>
  mutate(
    Revenue = factor(Revenue,
      levels = c(FALSE, TRUE),
      labels = c("No purchase", "Purchase")),
    Weekend = factor(Weekend),
    across(c(OperatingSystems, Browser,
             Region, TrafficType), factor)
  )
```

We checked variable types, missingness, class balance, and plausible ranges before modeling.

## Objective 1: Which model predicts purchases best? {.section-slide}

**Owner: Angela Qian**

Compare an interpretable baseline, a single decision tree, and a random forest using the same train/test split and preprocessing steps.

## We tested three classification approaches

| Method | Why include it? |
|---|---|
| Logistic regression | Clear baseline and interpretable coefficients |
| Decision tree | Simple nonlinear rules and interactions |
| Random forest | Ensemble that can capture complex patterns |

Regression-focused methods were not selected because the business outcome is categorical.

## The split protects the final evaluation

```{r}
#| echo: true
#| eval: false
shop_split <- initial_split(
  shop, prop = .80, strata = Revenue
)

shop_train <- training(shop_split)
shop_test  <- testing(shop_split)
```

- **Training set:** 80% of sessions
- **Test set:** 20% of sessions
- **Stratification:** preserves the 15.5% purchase rate

## One recipe keeps preprocessing consistent

```{r}
#| echo: true
#| eval: false
shop_rec <- recipe(Revenue ~ ., data = shop_train) |>
  step_novel(all_nominal_predictors()) |>
  step_unknown(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())
```

The same recipe was used for all three candidate models.

## The random forest ranked sessions best

```{r}
#| label: fig-model-comparison
#| fig-width: 10.5
#| fig-height: 4.5
model_metrics |>
  select(Model, `ROC AUC`, Recall, Precision) |>
  pivot_longer(-Model, names_to = "Metric", values_to = "Value") |>
  ggplot(aes(Value, Model, color = Metric)) +
  geom_point(size = 6) +
  scale_color_manual(values = c("#326B77", "#E9A23B", "#F28E5B")) +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top", panel.grid.major.y = element_blank())
```

The random forest achieved **`r number(test_auc, accuracy = .001)` ROC AUC** on unseen test data.

## A lower threshold finds more purchasers

```{r}
#| label: fig-thresholds
#| fig-width: 10
#| fig-height: 4.2
threshold_metrics |>
  pivot_longer(-Threshold, names_to = "Metric", values_to = "Value") |>
  ggplot(aes(Threshold, Value, fill = Metric)) +
  geom_col(position = position_dodge(.75), width = .68) +
  geom_text(aes(label = percent(Value, accuracy = .1)),
            position = position_dodge(.75), vjust = -.35,
            size = 5.7, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(labels = percent, limits = c(0, .9)) +
  labs(x = "Purchase probability threshold", y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top", panel.grid.major.x = element_blank())
```

At 30%, recall rises to 75.1% while precision remains 62.5%.

## Objective 1 finding {.finding-slide}

> Use the random forest for ranking sessions, then select the threshold based on outreach capacity and the cost of missed purchasers.

- Best test ROC AUC: 0.919
- A 30% threshold finds three out of four purchasing sessions
- A 50% threshold produces fewer but more precise alerts

## Objective 2: What behavior signals conversion? {.section-slide}

**Owner: Bhamini Pillai**

Explain which customer-journey variables matter and whether engagement patterns remain useful without the strongest page-value signal.

## Purchasers spend longer on product pages

```{r}
#| label: fig-duration-final
#| fig-width: 9.5
#| fig-height: 4.3
shop |>
  group_by(Revenue) |>
  summarise(Duration = median(ProductRelated_Duration), .groups = "drop") |>
  ggplot(aes(Revenue, Duration, fill = Revenue)) +
  geom_col(width = .58, show.legend = FALSE) +
  geom_text(aes(label = comma(Duration, accuracy = 1)),
            vjust = -.4, size = 6.2, fontface = "bold") +
  scale_fill_manual(values = c("#326B77", "#F28E5B")) +
  scale_y_continuous(limits = c(0, 1300), labels = comma) +
  labs(x = NULL, y = "Median product-page duration (seconds)") +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.x = element_blank())
```

Purchasing sessions had a median product-page duration of 1,110 seconds versus 510 seconds.

## Page value dominates model importance

```{r}
#| label: fig-importance
#| fig-width: 10
#| fig-height: 4.6
rf_importance |>
  slice_head(n = 8) |>
  ggplot(aes(Importance, reorder(Feature, Importance))) +
  geom_col(fill = "#326B77", width = .68) +
  labs(x = "Permutation importance", y = NULL) +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.y = element_blank())
```

Exit behavior, engagement, and administrative activity also contribute to prediction.

## We ran a stricter behavioral model

Page value is closely connected with transaction activity, so we also fitted logistic regression **without PageValues**.

```{r}
#| echo: true
#| eval: false
behavior_recipe <- recipe(Revenue ~ ., data = shop_train) |>
  step_rm(PageValues) |>
  step_dummy(all_nominal_predictors()) |>
  step_normalize(all_numeric_predictors())
```

This model asks whether earlier browsing behavior still contains predictive information.

## Behavior alone provides moderate ranking power

:::: {.columns}
::: {.column width="48%"}

### Without page value

- ROC AUC: **0.715**
- Accuracy: **83.9%**
- Recall at 50%: **3.7%**

:::
::: {.column width="48%"}

### Interpretation

Behavior helps rank sessions, but the default 50% cutoff misses most purchases.

The model needs threshold tuning or class-sensitive training before operational use.

:::
::::

## Longer product engagement raises predicted probability

```{r}
#| label: fig-duration-pdp
#| fig-width: 10
#| fig-height: 4.4
duration_pdp |>
  ggplot(aes(Duration, Probability)) +
  geom_line(color = "#326B77", linewidth = 1.7) +
  geom_point(color = "#F28E5B", size = 3.5) +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = percent) +
  labs(x = "Product-related duration (seconds)",
       y = "Average predicted purchase probability") +
  theme_minimal(base_size = 18)
```

The partial-dependence pattern is an association, not a causal effect.

## Objective 2 finding {.finding-slide}

> Page value is the strongest model signal, while product engagement and exit behavior provide earlier clues about conversion.

- Use page value carefully because it is close to the transaction outcome
- Monitor exit rate as a possible friction signal
- Treat long product browsing as an opportunity for assistance or remarketing

## Objective 3: Which segments deserve attention? {.section-slide}

**Owner: Willy Su**

Estimate how purchase probability changes across months and visitor types while the random forest holds the rest of the sample constant.

## Observed conversion peaks in November

```{r}
#| label: fig-observed-month-final
#| fig-width: 10.5
#| fig-height: 4.4
shop |>
  mutate(Month = factor(Month, levels = month_order)) |>
  group_by(Month) |>
  summarise(Rate = mean(Revenue == "Purchase"), .groups = "drop") |>
  ggplot(aes(Month, Rate, group = 1)) +
  geom_line(color = "#326B77", linewidth = 1.6) +
  geom_point(aes(color = Rate), size = 4.8) +
  scale_color_gradient(low = "#8EC6C5", high = "#F28E5B", guide = "none") +
  scale_y_continuous(labels = percent, limits = c(0, .30)) +
  labs(x = NULL, y = "Observed purchase rate") +
  theme_minimal(base_size = 18)
```

November's observed purchase rate was 25.4%.

## The segment analysis uses the ensemble model

For every month or visitor type, we:

1. Set that feature to one value for all test sessions
2. Predicted each session with the fitted random forest
3. Averaged the predicted purchase probabilities
4. Repeated the process for every category

This partial-dependence approach summarizes model behavior while holding the test population constant.

## Month remains important after adjustment

```{r}
#| label: fig-month-pdp-final
#| fig-width: 10.5
#| fig-height: 4.4
month_pdp |>
  ggplot(aes(Month, Probability, group = 1)) +
  geom_line(color = "#326B77", linewidth = 1.6) +
  geom_point(color = "#F28E5B", size = 4.8) +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Average predicted purchase probability") +
  theme_minimal(base_size = 18)
```

The model assigns the strongest adjusted probabilities to late-year sessions.

## Visitor type changes predicted conversion

```{r}
#| label: fig-visitor-pdp-final
#| fig-width: 9.5
#| fig-height: 4.3
visitor_pdp |>
  ggplot(aes(Probability, reorder(VisitorType, Probability))) +
  geom_col(fill = "#326B77", width = .62) +
  geom_text(aes(label = percent(Probability, accuracy = .1)),
            hjust = -0.2, fontface = "bold", size = 5.8) +
  scale_x_continuous(labels = percent, limits = c(0, .23)) +
  labs(x = "Average predicted purchase probability", y = NULL) +
  theme_minimal(base_size = 18) +
  theme(panel.grid.major.y = element_blank())
```

Visitor type should inform audience strategy, but the differences are smaller than the page-value signal.

## Objective 3 finding {.finding-slide}

> Timing and visitor type can improve campaign planning, but they should complement—not replace—session behavior.

- Prepare conversion campaigns before the late-year peak
- Test different messages for new and returning visitors
- Compare traffic quality within each segment
- Validate segment strategies through controlled experiments

## Three findings answer the business problem {.summary-slide}

1. **Prediction:** the random forest ranked purchase likelihood well, with 0.919 test ROC AUC.
2. **Behavior:** page value led the model; product engagement and exit behavior added useful signals.
3. **Segments:** late-year timing and visitor type changed predicted conversion opportunities.

## Recommended business actions {.recommend-slide}

:::: {.columns}
::: {.column width="31%"}

### Prioritize

Score sessions and use a threshold matched to available outreach capacity.

:::
::: {.column width="31%"}

### Improve

Review high-exit journeys and support visitors showing sustained product interest.

:::
::: {.column width="31%"}

### Test

Run controlled experiments on timing, audiences, and remarketing messages.

:::
::::

## Results have important limitations

- The data represents sessions, not a longitudinal customer history
- The source does not identify the retailer or product categories
- Page value may contain information close to the final transaction
- Class imbalance makes accuracy potentially misleading
- Model relationships do not prove that a business action will cause conversion
- Current performance should be validated on newer company data

## Next steps move from model to decision

1. Tune hyperparameters with cross-validation
2. Compare probability calibration and business costs
3. Validate the threshold using campaign capacity
4. Test explanations with SHAP values
5. Deploy a simple scoring app
6. Measure incremental conversion through an experiment

## Appendix {.appendix-slide}

::: {.appendix-card}
### Project resources

**GitHub repository**  
[View the Group 6 project repository](https://github.com/angelakew/ibm6540-final-report)

**Presentation slides**  
[Open the published RevealJS presentation](https://angelakew.github.io/ibm6540-final-report/)

**Project dataset**  
[View the UCI Online Shoppers dataset abstract](https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset)

**Dataset citation**  
[UCI dataset DOI: 10.24432/C5F88Q](https://doi.org/10.24432/C5F88Q)

**Model availability**  
The reproducible model analysis is included in the project QMD. A public scoring app is a planned next step.
:::

::: {.notes}
[Sources]
- UCI Machine Learning Repository: https://archive.ics.uci.edu/dataset/468/online+shoppers+purchasing+intention+dataset
- Dataset DOI: https://doi.org/10.24432/C5F88Q
:::

## Reproducibility notes

- Random seed: 6540
- Stratified 80/20 train/test split
- Shared preprocessing recipe across candidate models
- Test set used only for final comparison
- R packages: tidymodels, tidyverse, ranger, and Quarto
- Complete analysis code is contained in this QMD file

## The business value is focused learning {.statement-slide .closing-slide}

> Predict purchase likelihood, explain the customer journey, and test the actions that follow.

The model is a starting point for better decisions—not a substitute for experimentation.

### Project links

[GitHub repository](https://github.com/angelakew/ibm6540-final-report) · [Published presentation](https://angelakew.github.io/ibm6540-final-report/)
